% $Id: design_choices.tex,v 1.237 2003/10/24 10:24:24 mkomu Exp $

%%%%%%%%%%%%%%%%%%%% GENERAL TODO SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% - Merge Torsten's comments in this paper
%   - Long sections should be split into smaller
%     - Example: \textbf{Keyword} begins this section..
%   - Some sections contain repeated information, is this bad?
%     - Resolver architecture contains redundant information
%       - Network socket API: The motivation behind the flags...
%       - If both transparency and AI_HIP...
%     - Userspace network applications
%        - HIP aware applications may require...
%        - Technically this is a correct way...
% - Jouni Korhonen suggested
%   - more stuff to DNS part -> Miika's thesis corrects this.
%   - more punctuality with the solutions (state/sequence transition graphs)
%   - separate ``we did not implement x because'' things from the text
%     because they are being repeated all over the text
%   - clarify the things we did not implement
%   - clarify why we implemented x with method y
% - mention about transport/tunnel mode and which/why our implementation
%   uses it
% - Merge Henderson's email comments in this paper
%   - One packet buffer - compare to Jukka's thoughts
%   - HIs can be distributed using CERTs - requires PKI
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,titlepage]{article}
\usepackage{fancyhdr}
\usepackage{a4wide}
%\usepackage{alltt}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{supertabular}
\usepackage{array}          % XX TODO: is this needed?
%\usepackage{enumerate}
\usepackage{varioref}
\usepackage{hyperref}       % keep this as the last one
\usepackage{acronym}
\begin{document}

\input{design_choices_macros}

\title{Host Identity Protocol for Linux}
\author{
	Mika Kousa, \texttt{mika.kousa@hiit.fi} \\
        Miika Komu, \texttt{miika.komu@hiit.fi} \\
	Kristian Slavov, \texttt{kristian.slavov@hiit.fi} \\
        Catharina Candolin, \texttt{candolin@tml.hut.fi} \\
	Janne Lundberg, \texttt{jlu@tcs.hut.fi} \\
	\\
        Helsinki Institute for Information Technology \\       
        Helsinki University of Technology: \\
        Telecommunications Software and Multimedia Laboratory}
\maketitle


\fancyhf{} % Clear all fields
\fancyhead[C]{\small Host Identity Protocol for Linux}
\fancyfoot[C]{ \thepage }
\pagestyle{fancy}

\tableofcontents

\newpage

\section*{Acknowledgements}

This research is funded by the National Technology Agency of Finland
(TEKES), Elisa Communications, Ericsson, Nokia, TeliaSonera, Creanor,
and More Magic Software. We thank Jukka Ylitalo, Jorma Wall, Petri
Jokela, and especially Pekka Nikander from Ericsson Research for
fruitful cooperation and interoperability testing with their
implementation of HIP for FreeBSD. Furthermore, we are grateful for
the valuable discussions and comments we have had with Andrew
McGregor, Thomas Henderson, and Jeff Ahrenholz. Torsten R{\"u}ger
(HIIT) and Jouni Korhonen (TeliaSonera) gave also some valuable
comments on this document. Jaakko Kangasharju has kindly provided some
editorial comments.

\newpage

%%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

One of the main problems with IP has been its lack of security.
Although IPsec and DNSSEC have provided some level of security to IP,
the notion of a endpoint identity for hosts is still missing.
Typically, the IP address of the host has been used as the host
identity in the transport layer, regardless of the fact that it is
nothing more than routing information.  The purpose of the Host
Identity Protocol (HIP) architecture is to enhance the IP protocol
with a new cryptographically based namespace. The Host Identity serves
as the identity of the host at the transport layer, whereas the IP
address at the network layer is merely used for routing purposes.  The
separation of the transport and network layer identifiers facilitates
the support of end-host mobility and multihoming.  Further, the
cryptographic nature of the new namespace makes it possible to support
end-host mobility and multihoming in secure way.  In this document, we
briefly describe the HIP architecture, and present our implementation
of HIP for Linux.

\end{abstract}

\clearpage

\input{acronyms.tex}

\input{hipl.tex}

\bibliographystyle{plain}
\bibliography{design_choices}


\appendix
\newpage
\pagebreak

\include{catalogue}

\end{document}
